---
title: "KI - Project"
author: "Anna Sofie Christensen, Carrie Lovejoy, Sofie Schubert Elving"
date: '2022-12-30'
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)
library("survival")
library("truncreg")
```
The data set can be found here:
https://www.kaggle.com/datasets/reihanenamdari/breast-cancer?resource=download

## 1) The data and alterations

```{r}
bc <- read.csv("C:/Users/sofel/OneDrive/Dokumenter/AarhusUniversitet/3.semester/KI/R/Breast_Cancer.csv")
```

We alternate our data a bit:

We first change our Age variable to be age after 30:
```{r}
bc$Age <- bc$Age - 30
```

Then changing Status, Progesterone.Status and Estrogen.Status to dummy variables and renaming those:
```{r}
bc$Status <- ifelse(bc$Status == "Dead", 1, 0)
bc$Progesterone.Status <- ifelse(bc$Progesterone.Status == "Positive", 1, 0)
bc$Estrogen.Status <- ifelse(bc$Estrogen.Status == "Positive", 1, 0)
```

```{r}
bc <- bc %>%
  rename(
    Dead = "Status",
    Prog.Pos = "Progesterone.Status",
    Estro.Pos = "Estrogen.Status",
    Reg.N.Pos = "Reginol.Node.Positive",
    Reg.N.Ex = "Regional.Node.Examined"
  )
```


## 2) Interaction terms

We add an age^2 variable:
```{r}
bc$Agesq <- bc$Age^2
```


For Regional.Node.Examined and Regional.Node.Positive, we can't have more positive lymph nodes than we have examined. This can be seen in this plot:
```{r}
plot(bc$Reg.N.Ex, bc$Reg.N.Pos)
```


## 3) Used models
The three models we test on are given here.

### Model 1
Model of the data, leaving out variables that are linked (perfect multicollinearity)
$$
\begin{aligned}
    SurvivalMonths_i &= \beta_0 + \beta_1 Age_i + \beta_2 Race_i + \beta_3 MaritalStatus_i + \beta_4 Grade_i \\ 
    & \quad + \beta_5 AStage_i + \beta_6 TumorSize_i + \beta_7 EstroPos_i + \beta_8 ProgPos_i \\
    & \quad + \beta_9 RegNPos_i + \beta_{10} RegNEx_i + u_i
\end{aligned}
$$

### Model 2
Model 1 with interaction terms and MaritalStatus dropped
$$
\begin{aligned}
    SurvivalMonths_i &= \beta_0 + \beta_1 Age_i + \beta_2 Age^2_i + \beta_3 Race_i + \beta_4 Grade_i + \beta_5 AStage_i \\
    & \quad + \beta_6 TumorSize_i + \beta_7 EstroPos_i + \beta_8 ProgPos_i  \\
    & \quad +  \beta_9 (EstroPos_i \cdot ProgPos_i) \\ \nonumber
    & \quad + \beta_{10} RegNPos_i + \beta_{11} (RegNPos_i \cdot RegNEx_i) + u_i
\end{aligned}
$$

### Model 3
Model 2 without the EstroPos and ProgPos interaction term.
$$
\begin{aligned}
    \nonumber SurvivalMonths_i &= \beta_0 + \beta_1 Age_i + \beta_2 Age^2_i + \beta_3 Race_i + \beta_4 Grade_i + \beta_5 AStage_i \\ 
    & \quad + \beta_6 TumorSize_i + \beta_7 EstroPos_i + \beta_8 ProgPos_i + \beta_9 RegNPos_i \\ \nonumber
    & \quad + \beta_{10} (RegNPos_i \cdot RegNEx_i) + u_i
\end{aligned}
$$

## 4) Naive OLS

We run a naive OLS on model 1:
```{r}
LM1 <-
  lm(
    data = bc,
    formula = Survival.Months ~ Age + Race + Marital.Status + Grade + A.Stage + Tumor.Size + Estro.Pos + Prog.Pos + Reg.N.Pos + Reg.N.Ex
  )
summary(LM1)
```

And on model 2:
```{r}
LM2 <-
  lm(
    data = bc,
    formula = Survival.Months ~ Age + Agesq + Race + Grade + A.Stage + Tumor.Size + Estro.Pos *
      Prog.Pos + Reg.N.Pos + Reg.N.Ex:Reg.N.Pos
  )
summary(LM2)
```

## 5) Truncation

We use the truncreg() function to test our regression taking truncation into consideration.

We first do this on model 1:
```{r}
trunc1 <-
  truncreg(
    data = bc,
    formula = Survival.Months ~ Age + Race + Marital.Status + Grade + A.Stage + Tumor.Size + Estro.Pos + Prog.Pos + Reg.N.Pos + Reg.N.Ex,
    point = 1.0,
    direction = "left"
  )
summary(trunc1)
```

Then on model 2:
```{r}
trunc2 <-
  truncreg(
    data = bc,
    formula = Survival.Months ~ Age + Agesq + Race + Grade + A.Stage + Tumor.Size + Estro.Pos *
      Prog.Pos + Reg.N.Pos + Reg.N.Ex:Reg.N.Pos,
    point = 1.0,
    direction = "left"
  )
summary(trunc2)
```


## 6) Censoring
We use the survreg()-function to take our right sided censoring into consideration

First on model 1:
```{r}
tobit1 <-
  survreg(
    data = bc,
    formula = Surv(Survival.Months, Dead, type = 'right') ~ Age + Race + Marital.Status + Grade + A.Stage + Tumor.Size + Estro.Pos + Prog.Pos + Reg.N.Pos + Reg.N.Ex,
    dist = "gaussian",
    x = TRUE
  )
summary(tobit1)
```

Then on model 2:
```{r}
tobit2 <-
  survreg(
    data = bc,
    formula = Surv(Survival.Months, Dead, type = 'right') ~ Age + Agesq + Race + Grade + A.Stage + Tumor.Size + Estro.Pos *
      Prog.Pos + Reg.N.Pos + Reg.N.Ex:Reg.N.Pos,
    dist = "gaussian"
  )
summary(tobit2)
```

Finally on model 3:
```{r}
tobit3 <-
  survreg(
    data = bc,
    formula = Surv(Survival.Months, Dead, type = 'right') ~ Age + Agesq + Race + Grade + A.Stage + Tumor.Size + Estro.Pos + Prog.Pos + Reg.N.Pos + Reg.N.Pos:Reg.N.Ex,
    dist = "gaussian",
    x = TRUE
  )
summary(tobit3)
```

### Adjustment factors

We take a look at our tobit3 model after applying the adjustment factors.

```{r}
adj = pnorm(abs(apply(tobit3$x,2,FUN=mean) * tobit3$coef) / sqrt(diag(vcov(tobit3)))[-1])
adj
```

```{r}
adj * tobit3$coef
```


## 7) Figures

### Error terms
We check if our error terms are normally distributed:
```{r}
res = residuals(tobit3)
qqnorm(res, main = "Normal QQ-plot of Residuals")
qqline(res)
```


### Homoskedasticity 
```{r}
plot(fitted(tobit3),
     res,
     main = "Residuals vs Fitted",
     xlab = "Fitted values",
     ylab = "Residuals")
abline(lm(res ~ fitted(tobit3)), col = "red")
abline(0, 0)
```
